---
title: "Taking a look at dilution ratios in cocktails"

date: "`r Sys.Date()`"
author:
- Harold Pimentel

output:
  html_document:
    toc: false
    theme: default
    fig_height: 7
    fig_width: 11
---

Looking at dilution ratios and cocktails is a function of the alcohol by volume.

The data below is from the book [_Liquid Intelligence_](https://www.amazon.com/Liquid-Intelligence-Science-Perfect-Cocktail/dp/0393089037) by Dave Arnold.

# setting up the environment

Required packages:

```{r}
library('dplyr')
```

Code to fit a second order polynomial.

```{r}
polynomial_fit <- function(df, which_method) {
  df <- dplyr::filter(df, method == which_method)
  model <- lm(dilution_ratio ~ poly(start_abv, 2, raw = TRUE), data = df)
  df <- mutate(df, predicted = predict(model, df))
  plot(df$start_abv, df$dilution_ratio)
  lines(df$start_abv, df$predicted)
  list(data = df, model = model)
}
```

Load the data.

```{r}
profiles <- read.csv('../data/profile.csv', header = TRUE, stringsAsFactors = FALSE)
profiles <- mutate(profiles, start_abv = start_abv / 100, finish_abv = finish_abv / 100)
profiles <- mutate(profiles, dilution_ratio = (finished_volume - mix_volume) / mix_volume)
```

Fit the model using the stirred data.

```{r}
stirred <- polynomial_fit(profiles, 'stirred')
summary(stirred$model)
```

For the model using the shaken data.

```{r}
shaken <- polynomial_fit(profiles, 'shaken')
summary(shaken$model)
```

# Taking a look at the results

Using the model from the book, the `predicted_book` dilution ratio is too large resulting in a very large `predicted_book_finish`, the predicted finished volume (using the books model).

```{r}
shaken_model_book <- function(abv) 1.567 * abv * abv + 1.742 * abv + 0.203
shaken$data %>%
  select(name, start_abv, mix_volume, finished_volume, dilution_ratio, predicted) %>%
  mutate(
    predicted_finish = mix_volume + mix_volume * predicted,
    predicted_book = shaken_model_book(start_abv),
    predicted_book_finish = mix_volume + mix_volume * predicted_book
  ) %>%
  kable(digits = 2)
```

Making the second order coefficient negative seems to fix the problem:

```{r}
shaken_model_book <- function(abv) -1.567 * abv * abv + 1.742 * abv + 0.203
shaken$data %>%
  select(name, start_abv, mix_volume, finished_volume, dilution_ratio, predicted) %>%
  mutate(
    predicted_finish = mix_volume + mix_volume * predicted,
    predicted_book = shaken_model_book(start_abv),
    predicted_book_finish = mix_volume + mix_volume * predicted_book
  ) %>%
  kable(digits = 2)
```

Stirred data seems okay.

```{r}
stirred_model_book <- function(abv) -1.21 * abv * abv + 1.246 * abv + 0.145
stirred$data %>%
  select(name, start_abv, mix_volume, finished_volume, dilution_ratio, predicted) %>%
  mutate(
    predicted_finish = mix_volume + mix_volume * predicted,
    predicted_book = stirred_model_book(start_abv),
    predicted_book_finish = mix_volume + mix_volume * predicted_book
  ) %>%
  kable(digits = 2)
```
